---
title: "Week 2 — R Programming Fundamentals"
author: "Your Name"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Objects

Run this chunk and observe the Environment pane (top right) as each object is created.

```{r objects-demo}
# Assign values to objects
x <- 3 + 5
x <- x - 4
y <- x + 1

print(x)
print(y)
```

**Question:** What are the values of `x` and `y`? Did the result match your prediction from the reading?

*Your answer here:*

---

### 🔨 Exercise: Objects

```{r objects-exercise}
# 1. Create an object called bill_length_mm with the value 45.1
# Your code here

# 2. Convert it to centimetres and store in bill_length_cm
# Your code here

# 3. Print both
# Your code here
```

---

## Conditions

```{r conditions-demo}
# Basic if/else
grade <- 85

if (grade >= 90) {
  print("A")
} else if (grade >= 80) {
  print("B")
} else if (grade >= 70) {
  print("C")
} else {
  print("Below passing")
}
```

```{r booleans-demo}
# Boolean expressions
x <- 1 == 1    # TRUE or FALSE?
y <- 2 < 1     # TRUE or FALSE?
print(x)
print(y)
```

---

### 🔨 Exercise: Conditions

```{r conditions-exercise}
# Write an if/else that prints "Long bill" if bill_length_mm > 43, "Short bill" otherwise
bill_length_mm <- 47.5

# Your code here
```

---

## Loops

```{r loops-demo}
# A for loop evaluating each grade
grades <- c(78, 92, 65, 88, 71)

for (grade in grades) {
  if (grade >= 70) {
    print(paste(grade, "— Pass"))
  } else {
    print(paste(grade, "— Fail"))
  }
}
```

---

### 🔨 Exercise: Loops

```{r loops-exercise}
# Bill lengths for 6 penguins
bill_lengths <- c(39.1, 45.2, 36.7, 50.3, 42.8, 31.1)

# Write a for loop that prints "Long" if > 43, "Short" otherwise
# Your code here
```

---

## Functions

```{r functions-demo}
# Define a function
classify_bill <- function(bill_length_mm) {
  if (bill_length_mm > 43) {
    return("Long")
  } else {
    return("Short")
  }
}

# Call the function
classify_bill(39.1)
classify_bill(47.5)
```

---

### 🔨 Exercise: Functions

```{r functions-exercise}
# Write a function called evaluate_grade that takes a grade argument
# and returns "Pass" if >= 70, "Fail" otherwise

# Your function here


# Test it:
# evaluate_grade(85)
# evaluate_grade(62)
```

---

## Vectors and Data Frames

```{r vectors-demo}
# Vectors
species <- c("Adelie", "Chinstrap", "Gentoo")
bill_lengths <- c(39.1, 46.5, 47.3)

# Operate on all elements at once
bill_lengths_cm <- bill_lengths / 10
print(bill_lengths_cm)
```

```{r dataframe-demo}
# Create a data frame
penguin_df <- data.frame(
  species      = species,
  bill_length  = bill_lengths,
  bill_cm      = bill_lengths_cm
)

# Access a column
penguin_df$species
```

---

### 🔨 Exercise: Grade Calculator (put it all together)

```{r grade-calculator}
# Stage 1: Create a vector of 5 student grades
grades <- c(80, 95, 65, 72, 88)

# Stage 2: Add 5 bonus points to every grade
# Your code here

# Stage 3: Write an if/else for one grade (pass/fail at 70)
# Your code here

# Stage 4: Wrap Stage 3 in a for loop over all grades
# Your code here

# Stage 5: Wrap Stage 4 in a function called evaluate_grades()
#           that takes a vector as input
# Your code here

# Call your function:
# evaluate_grades(grades)
```
