---
title: "Week 5 — Inferential Statistics and Hypothesis Testing"
author: "Your Name"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, message=FALSE}
library(tidyverse)
library(palmerpenguins)
```

## Point Estimate

```{r point-estimate}
# Mean bill length as a point estimate for the population mean
point_est <- mean(penguins$bill_length_mm, na.rm = TRUE)
cat("Point estimate (mean bill length):", round(point_est, 2), "mm\n")
```

---

## Standard Error via Bootstrapping

```{r bootstrap-se}
set.seed(42)

# Draw 300 resamples and compute the mean for each
boot_means <- replicate(300, {
  resample <- penguins %>%
    drop_na(bill_length_mm) %>%
    sample_n(size = nrow(drop_na(penguins, bill_length_mm)), replace = TRUE)
  mean(resample$bill_length_mm)
})

se <- sd(boot_means)
cat("Standard error:", round(se, 4), "\n")

# The bootstrap distribution should be approximately normal
hist(boot_means, main = "Bootstrap distribution of mean bill length",
     xlab = "Mean bill length (mm)")
```

---

## 95% Confidence Interval

```{r confidence-interval}
lower <- point_est - 1.96 * se
upper <- point_est + 1.96 * se

cat("95% CI:", round(lower, 2), "to", round(upper, 2), "\n")
```

**Write a plain-language interpretation of this confidence interval:**

*Your interpretation here:*

---

## Normality Check

```{r normality-check}
# Check normality for Gentoo flipper lengths
gentoo <- penguins %>%
  filter(species == "Gentoo") %>%
  drop_na(flipper_length_mm)

# Histogram
ggplot(gentoo, aes(x = flipper_length_mm)) +
  geom_histogram(bins = 20) +
  labs(title = "Gentoo flipper length distribution")

# Shapiro-Wilk test
shapiro.test(gentoo$flipper_length_mm)
```

**Question:** Based on the histogram and Shapiro-Wilk result, is the normality assumption met for Gentoo flipper lengths?

*Your answer here:*

---

## Comparing Two Groups

```{r two-group-test}
# Compare flipper length between male and female Gentoo penguins
gentoo_sex <- penguins %>%
  filter(species == "Gentoo") %>%
  drop_na(sex, flipper_length_mm)

# t-test (parametric)
t.test(flipper_length_mm ~ sex, data = gentoo_sex)
```

**Write the full hypothesis test conclusion:**

*H₀:*  
*Hₐ:*  
*Test used and why:*  
*p-value:*  
*Conclusion:*  
*Possible error type:*

---

## Chi-Squared Test

```{r chisq-test}
# Is species independent of sex?
chisq.test(table(penguins$species, penguins$sex))
```

---

## 🔨 Exercise: ANOVA

```{r anova-exercise}
# 1. Run a one-way ANOVA to test whether mean body_mass_g differs across species
# Your code here

# 2. If significant, run pairwise.t.test() with Bonferroni correction
# Your code here

# 3. Which pairs of species differ significantly?
# Your answer as a comment here
```
