---
title: "Week 7 — Survival Analysis"
author: "Your Name"
---

```{r setup, message=FALSE}
library(tidyverse)
library(survival)
```

## Data

This data was obtained from the publicly available Health and Retirement Study: https://hrs.isr.umich.edu/about

```{r load-data}
surv_df <- read_csv("hrs/hrs_analyze.csv") %>%
  drop_na(mos_to_death, died, female, age_2014, bp_dx, depress_dx)

glimpse(surv_df)
```

---

## Data check: censoring

```{r censoring-check}
# What proportion are censored?
surv_df %>%
  count(died) %>%
  mutate(prop = n / sum(n))
```

**Question:** What does it mean that some observations have `died = 0`? Should we remove them?

*Your answer here:*

---

## Kaplan-Meier Curve

```{r km-fit}
# Fit KM curve by sex
km_fit <- survfit(Surv(mos_to_death, died) ~ female, data = surv_df)
summary(km_fit, times = c(12, 24, 36, 48))
```

```{r km-plot}
# Plot
plot(km_fit,
     col        = c("steelblue", "coral"),
     lwd        = 2,
     xlab       = "Months",
     ylab       = "Survival probability",
     main       = "Kaplan-Meier curves by sex")
legend("topright",
       legend = c("Male", "Female"),
       col    = c("steelblue", "coral"),
       lwd    = 2)
```

**Interpret the KM curves in 1–2 sentences using descriptive language:**

*Your interpretation here ("In this sample..."):*

---

## Log Rank Test

```{r logrank-test}
survdiff(Surv(mos_to_death, died) ~ female, data = surv_df)
```

**State the full hypothesis test result:**

*H₀:*  
*Hₐ:*  
*p-value:*  
*Conclusion:*

---

## Cox Proportional Hazards Regression

```{r cox-model}
cox_model <- coxph(
  Surv(mos_to_death, died) ~ female + age_2014 + bp_dx + depress_dx,
  data = surv_df
)
summary(cox_model)
```

```{r hazard-ratios}
# Extract hazard ratios
exp(coef(cox_model))
```

**Interpret the hazard ratio for `depress_dx` in plain language:**

*Your interpretation here (include: predictor, outcome, HR value, direction, significance):*

---

## 🔨 Exercise: Your own KM curve

```{r km-exercise}
# 1. Create a binary age group: "younger" (age_2014 < 65) and "older" (age_2014 >= 65)
surv_df <- surv_df %>%
  mutate(age_group = ifelse(age_2014 < 65, "younger", "older"))

# 2. Fit and plot a KM curve stratified by age_group
# Your code here

# 3. Run a log rank test to compare survival by age_group
# Your code here

# 4. Interpret the result in plain language (as a comment)
```
