---
title: "Week 7 — Survival Analysis"
author: "Your Name"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, message=FALSE}
library(tidyverse)
library(survival)
```

> **Note:** Survival analysis requires time-to-event data with a censoring indicator.  
> This notebook uses a simulated dataset so no additional download is needed.  
> In class and in the lesson slides, the HRS dataset is used — replace below if you have access.

---

## Simulated survival data

```{r simulate-data}
set.seed(42)
n <- 200

surv_df <- tibble(
  id          = 1:n,
  # Time to event or censoring (months)
  months      = round(rexp(n, rate = 0.03) + 1),
  # Event status: 1 = event occurred, 0 = censored
  died        = rbinom(n, 1, prob = 0.6),
  # Group variable
  treatment   = sample(c("control", "treatment"), n, replace = TRUE),
  # Continuous covariate
  age         = round(rnorm(n, mean = 65, sd = 10))
)

# Cap time at 60 months (end of study)
surv_df <- surv_df %>%
  mutate(
    died   = ifelse(months > 60, 0, died),
    months = pmin(months, 60)
  )

glimpse(surv_df)
```

---

## Data check: censoring

```{r censoring-check}
# What proportion are censored?
surv_df %>%
  count(died) %>%
  mutate(prop = n / sum(n))
```

**Question:** What does it mean that some observations have `died = 0`? Should we remove them?

*Your answer here:*

---

## Kaplan-Meier Curve

```{r km-fit}
# Fit KM curve by treatment group
km_fit <- survfit(Surv(months, died) ~ treatment, data = surv_df)
summary(km_fit, times = c(12, 24, 36, 48))
```

```{r km-plot}
# Plot
plot(km_fit,
     col        = c("steelblue", "coral"),
     lwd        = 2,
     xlab       = "Months",
     ylab       = "Survival probability",
     main       = "Kaplan-Meier curves by treatment group")
legend("topright",
       legend = c("Control", "Treatment"),
       col    = c("steelblue", "coral"),
       lwd    = 2)
```

**Interpret the KM curves in 1–2 sentences using descriptive language:**

*Your interpretation here ("In this sample..."):*

---

## Log Rank Test

```{r logrank-test}
survdiff(Surv(months, died) ~ treatment, data = surv_df)
```

**State the full hypothesis test result:**

*H₀:*  
*Hₐ:*  
*p-value:*  
*Conclusion:*

---

## Cox Proportional Hazards Regression

```{r cox-model}
cox_model <- coxph(
  Surv(months, died) ~ treatment + age,
  data = surv_df
)
summary(cox_model)
```

```{r hazard-ratios}
# Extract hazard ratios
exp(coef(cox_model))
```

**Interpret the hazard ratio for `treatment` in plain language:**

*Your interpretation here (include: predictor, outcome, HR value, direction, significance):*

---

## 🔨 Exercise: Your own KM curve

```{r km-exercise}
# 1. Create a binary age group: "younger" (age < 65) and "older" (age >= 65)
surv_df <- surv_df %>%
  mutate(age_group = ifelse(age < 65, "younger", "older"))

# 2. Fit and plot a KM curve stratified by age_group
# Your code here

# 3. Run a log rank test to compare survival by age_group
# Your code here

# 4. Interpret the result in plain language (as a comment)
```
