---
title: "Week 6 — Linear and Logistic Regression"
author: "Your Name"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, message=FALSE}
library(tidyverse)
library(palmerpenguins)
```

> **Note:** The lesson slides use the HRS (Health and Retirement Study) dataset.  
> This notebook uses the Palmer Penguins dataset so no additional data download is needed.  
> Replace with your own dataset for your final project.

---

## Linear Regression

### Data check — is the outcome continuous and roughly normal?

```{r outcome-check}
# We'll predict body_mass_g from flipper_length_mm
penguins %>%
  drop_na(body_mass_g) %>%
  ggplot(aes(x = body_mass_g)) +
    geom_histogram(bins = 30) +
    labs(title = "Distribution of body mass (outcome variable)")
```

**Question:** Does `body_mass_g` look approximately normal? Is linear regression appropriate?

*Your answer here:*

---

### Fit the model

```{r linear-model}
penguins_clean <- penguins %>% drop_na(body_mass_g, flipper_length_mm, sex, species)

model_linear <- glm(body_mass_g ~ flipper_length_mm, data = penguins_clean)
summary(model_linear)
```

**Interpret the `flipper_length_mm` coefficient in plain language:**

*Your interpretation here (include: predictor, outcome, direction, magnitude, significance):*

---

### Multivariable linear regression

```{r multivariable-linear}
model_multi <- glm(body_mass_g ~ flipper_length_mm + species + sex, data = penguins_clean)
summary(model_multi)
```

**Pick one coefficient and interpret it (include "holding other variables constant"):**

*Your interpretation here:*

---

### Residual diagnostics

```{r residuals}
par(mfrow = c(1, 2))
plot(model_multi, which = 1)   # Residuals vs. Fitted
plot(model_multi, which = 2)   # QQ plot
```

**Question:** Do the residuals look well-behaved? Describe what you see.

*Your answer here:*

---

## Logistic Regression

```{r logistic-setup}
# Create a binary outcome: is the penguin above median mass?
median_mass <- median(penguins_clean$body_mass_g)
penguins_clean <- penguins_clean %>%
  mutate(heavy = as.integer(body_mass_g > median_mass))

table(penguins_clean$heavy)
```

```{r logistic-model}
model_logistic <- glm(heavy ~ flipper_length_mm + species,
                      family = binomial, data = penguins_clean)
summary(model_logistic)
```

```{r odds-ratios}
# Convert log-odds to odds ratios
exp(coef(model_logistic))
```

**Interpret the odds ratio for `flipper_length_mm`:**

*Your interpretation here (include: predictor, outcome, OR value, direction, significance):*

---

## 🔨 Exercise: Multicollinearity Check

```{r vif-exercise}
# Install car if needed: install.packages("car")
library(car)

# Check VIF for the multivariable linear model
vif(model_multi)

# Question: Are any VIF values concerning? What does this tell you?
# Your answer here as a comment:
```
