---
title: "Week 6 — Linear and Logistic Regression"
author: "Your Name"
---

```{r setup, message=FALSE}
library(tidyverse)
```

## Data

This data was obtained from the publicly available Health and Retirement Study: https://hrs.isr.umich.edu/about

```{r load-data}
hrs_df <- read_csv("hrs/hrs_analyze.csv")
glimpse(hrs_df)
```

---

## Linear Regression

### Data check — is the outcome continuous and roughly normal?

```{r outcome-check}
# We'll predict systolic blood pressure (bp_sys) from diastolic blood pressure (bp_dia)
hrs_df %>%
  drop_na(bp_sys) %>%
  ggplot(aes(x = bp_sys)) +
    geom_histogram(bins = 30) +
    labs(title = "Distribution of systolic blood pressure (outcome variable)")
```

**Question:** Does `bp_sys` look approximately normal? Is linear regression appropriate?

*Your answer here:*

---

### Fit the model

```{r linear-model}
hrs_clean <- hrs_df %>% drop_na(bp_sys, bp_dia, age_2014, female, depress_dx, bp_dx)

model_linear <- glm(bp_sys ~ bp_dia, data = hrs_clean)
summary(model_linear)
```

**Interpret the `bp_dia` coefficient in plain language:**

*Your interpretation here (include: predictor, outcome, direction, magnitude, significance):*

---

### Multivariable linear regression

```{r multivariable-linear}
model_multi <- glm(bp_sys ~ bp_dia + age_2014 + female, data = hrs_clean)
summary(model_multi)
```

**Pick one coefficient and interpret it (include "holding other variables constant"):**

*Your interpretation here:*

---

### Residual diagnostics

```{r residuals}
par(mfrow = c(1, 2))
plot(model_multi, which = 1)   # Residuals vs. Fitted
plot(model_multi, which = 2)   # QQ plot
```

**Question:** Do the residuals look well-behaved? Describe what you see.

*Your answer here:*

---

## Logistic Regression

```{r logistic-setup}
# Depression diagnosis (depress_dx) is already a binary 0/1 outcome
table(hrs_clean$depress_dx)
```

```{r logistic-model}
# Predict depression diagnosis from hypertension diagnosis
model_logistic <- glm(depress_dx ~ bp_dx, family = binomial, data = hrs_clean)
summary(model_logistic)
```

```{r odds-ratios}
# Convert log-odds to odds ratios
exp(coef(model_logistic))
```

**Interpret the odds ratio for `bp_dx`:**

*Your interpretation here (include: predictor, outcome, OR value, direction, significance):*

---

## 🔨 Exercise: Multicollinearity Check

```{r vif-exercise}
# Install car if needed: install.packages("car")
library(car)

# Check VIF for the multivariable linear model
vif(model_multi)

# Question: Are any VIF values concerning? What does this tell you?
# Your answer here as a comment:
```
