---
title: "Week 4 — Descriptive Statistics and Visualization"
author: "Your Name"
---

```{r setup, message=FALSE}
library(tidyverse)
library(palmerpenguins)
```

## Descriptive vs. Inferential — the key distinction

```{r penguin-summary}
# Compute mean bill length by species — this is descriptive
penguins %>%
  group_by(species) %>%
  summarize(mean_bill_mm = mean(bill_length_mm, na.rm = TRUE))
```

**Question:** Write one accurate and one inaccurate conclusion from the table above.

*Accurate (descriptive):*  
*Inaccurate (inferential overreach):*

---

## Grammar of Graphics

```{r ggplot-layers}
# Build a plot layer by layer

# Layer 1: data + coordinate system
penguins %>%
  ggplot()

# Layer 2: add a geometry
penguins %>%
  ggplot() +
    geom_histogram(mapping = aes(x = bill_length_mm))

# Layer 3: add color by species
penguins %>%
  ggplot() +
    geom_histogram(mapping = aes(x = bill_length_mm, fill = species),
                   bins = 30, alpha = 0.7)
```

### 🔨 Exercise: Grammar of Graphics

```{r ggplot-exercise}
# 1. Create a histogram of flipper_length_mm
# Your code here

# 2. Create a scatter plot of bill_length_mm (x) vs body_mass_g (y), colored by species
# Your code here

# 3. Add a title to your scatter plot using + labs(title = "...")
# Your code here
```

---

## Categorical Variables

```{r bar-plot}
# Bar plot of species counts
penguins %>%
  ggplot() +
    geom_bar(mapping = aes(x = species))
```

```{r factor-ordering}
# Control category order with factors
penguins %>%
  mutate(species = factor(species, levels = c("Gentoo", "Adelie", "Chinstrap"))) %>%
  ggplot() +
    geom_bar(mapping = aes(x = species))
```

### 🔨 Exercise: Categorical

```{r categorical-exercise}
# 1. Create a bar plot of penguin counts by island
# Your code here

# 2. Reorder the bars so they appear as: Biscoe, Dream, Torgersen
# Your code here

# 3. Compute the count and proportion of penguins per species
#    (proportion = count / sum(count))
# Your code here
```

---

## Continuous Variables

```{r histogram-demo}
# Always plot first to assess shape
penguins %>%
  ggplot() +
    geom_histogram(aes(x = body_mass_g), bins = 30)
```

**Question:** Does body_mass_g look normally distributed? Which summary statistics are appropriate?

*Your answer here:*

```{r summary-stats}
# Summary statistics
mean(penguins$body_mass_g, na.rm = TRUE)
sd(penguins$body_mass_g, na.rm = TRUE)
summary(penguins$body_mass_g)
```

```{r facet-histogram}
# Faceted by species — does shape differ by group?
penguins %>%
  ggplot() +
    geom_histogram(aes(x = body_mass_g), bins = 20) +
    facet_wrap(~species)
```

### 🔨 Exercise: Continuous

```{r continuous-exercise}
# 1. Plot a histogram of body_mass_g. Does it look normal?
# Your code here

# 2. Based on the shape, compute the appropriate summary statistics
# Your code here

# 3. Create a faceted histogram of body_mass_g by species
# Your code here
```

---

## EDA Gallery — Your Turn

```{r eda-your-choice}
# Create one additional plot of your choice that shows something interesting
# about the penguins dataset. Write a 1-2 sentence description below.

# Your plot here


# Description (use sample-appropriate language):
# "In this sample of 344 penguins..."
```
