---
title: "Week 1 — Setup and Orientation"
author: "Your Name"
date: "`r Sys.Date()`"
output: html_document
---

## Welcome

This notebook confirms your R and RStudio installation is working correctly and gives you your first taste of running R code.

---

## Step 1: Check your R version

Run the chunk below. You should see something like `"R version 4.4.x (...)"`.

```{r r-version}
R.version.string
```

---

## Step 2: Load the tidyverse

```{r load-tidyverse, message=FALSE}
library(tidyverse)
```

If this runs without a red error message, tidyverse is installed correctly.

---

## Step 3: Confirm tidyverse packages

```{r tidyverse-packages}
tidyverse_packages()
```

---

## Step 4: Your first R calculations

Run each chunk and observe the output.

```{r arithmetic}
# R as a calculator
2 + 2
sqrt(144)
10^3
```

```{r objects}
# Store a value in an object
my_name <- "Your Name Here"
my_name
```

---

## Step 5: Explore the Help pane

In the Console (not here), type `?mean` and press Enter. The Help pane will open with documentation for the `mean()` function.

Then try:

```{r help-example}
# Using a function we looked up
example_values <- c(10, 20, 30, 40, 50)
mean(example_values)
```

---

## ✅ Completion check

If all chunks above ran without error, your setup is complete.

**Write your name and today's date here as a completion note:**

*Name:*  
*Date:*  
*Any issues encountered:*
